---
title: "CCFM Example Application"
author: "Carly Hansen"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{CCFM_Example}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
## Setup

Load necessary libraries, and read in data and format by subsetting a specificed model into future and baseline periods. (Note that periods for the baseline and the observed dataset should be the same).

```{r}
#Load Libraries
library(CCFMR)
library(lubridate)

#Read in data
slcbcca=slc_bcca #Modeled data with time series of Date and model columns
slc_obs=slc_obs #Observed data

#Limit data to specified time period for a specified model
baseperiod=c(1955,1984) # define this based on available observed data
futureperiod=c(2055,2084) # define this based on your time period of interest and available modeled data
base=formatperiods(slcbcca,"ccsm4.1.rcp45",baseperiod)
future=formatperiods(slcbcca,"ccsm4.1.rcp45",futureperiod)

obs=formatperiods(slc_obs,"ObservedPrecip",baseperiod) # Use this as the base precipitation that will be scaled to a future period

```

Calculate the additive and multiplicative change factors using the average daily values.
```{r}
#Find average precipitation volume for each day of the year
futureavg=avgdaily(future)
baseavg=avgdaily(base)
#Calculate additive and multiplicative change factors
add=calccf(baseavg,futureavg,cftype="additive") #futureavg-baseavg
mult=calccf(baseavg,futureavg,cftype="multiplicative") #futureavg/baseavg
```

Calculate percentiles for the base modeled, future modeled, and base observed periods.
```{r}
#Calculate the percentiles (ECDF) of modeled values
base=calcecdf(base,"Precip")
future=calcecdf(future,"Precip")

#Calculate the percentiles (ECDF) of the observed values
obs=calcecdf(obs,"Precip")

#Visualize Results by plotting precip for the various records against the percentiles
plot(x=base$Precip,y=base$percentile,ylab='Percentile',xlab='Precipitation (mm/day)',main='ECDF',col=2)
points(x=future$Precip,y=future$percentile,col=4)
points(x=obs$Precip,y=obs$percentile,col=3)
legend(20,60,legend=c("Base Modeled","Future Modeled","Base Observed"),col=c(2,4,3),pch=c(1,1,1))

```

# Traditional Change Factor Methodology
Apply the change factors to the historical observations from the same time period as the baseline period.
```{r}
#Calculate the day of the year for observed data
obs$DOY=yday(obs$Date)

#Apply corresponding cf to observed dataset 
scaledadd=scalesingle(add,obs,cftype="additive")
scaledmult=scalesingle(mult,obs,cftype="multiplicative")
```


# Combined Change Factor Methodology
Create Difference and Ratio Plots:
Following the suggestion of Anandhi, et al., use ratios and difference plots to determine which type of change factor might be the most appropriate (if there is a curve, large slope, or large variation over a range in one CF type, use the other type. If there is no difference, use default of the multiplicative).

```{r,fig.show='hold'}
#Create a data frame to store the percentiles
percent=data.frame(percentlow=seq(1,100)) 

for (i in seq(0,99)){ # add values for each percentile (from 1-100) for both the base and future records
    percent$base[i+1]=mean(base[(base$percentlow==i),"Precip"])
    percent$future[i+1]=mean(future[(future$percentlow==i),"Precip"])
}
percent$base[is.nan(percent$base)]=0
percent$future[is.nan(percent$future)]=0

#Calculate ratio and differences for each bin
diff=percent$future-percent$base
ratio=percent$future/percent$base

plot(x=diff,y=seq(1,100),xlab='Difference of Future-Baseline',ylab='Percentile')
plot(x=ratio,y=seq(1,100),xlab='Ratio of Future/Baseline',ylab='Percentile')
```


Apply the change factor type for different ranges, using difference/ratio plots as a guide, and set a threshold for when the near-maxima values should use the additive change factor (regardless of difference/ratio plot results, which may not be very helpful near the maxima).
```{r}

#Format a dataframe for the combined change factor method (CCFM)
scaledccfm=merge(obs,add,by="DOY")
scaledccfm=merge(scaledccfm,mult,by="DOY")
names(scaledccfm)=c("DOY","Date","Precip","percentile","percentlow","addcf","multcf")

#Apply the single change factors over the specified ranges
# Example: scaledccfms=ccfm(scaledccfms,"Precip","addcf","multcf",70,95) # This would apply the multiplicative cf as default, then from 0-60 percentiles, additive would be applied, from 70-95, it would be left as multiplicative, and then above 95th percentile, it would be additive again. addcf and multcf are the columns that show the change factors at the respective DOY.
scaledccfm=ccfm(scaledccfm,"Precip","addcf","multcf",70,95)

```

# Summary of scaled and unscaled precip produced by different methods

Calculate Summary Statistics to Compare Different Scaling Methods to each other and to historical precipitation
```{r}
#Historical Observed
precipsummary(data=obs,precipcol="Precip")

#Scaled results produced with Combined Multiple Change Factors
precipsummary(scaledccfm,"scaled")

#Unscaled BCCA
precipsummary(future,"Precip")
#-----------------------------------------------------------
#Compare the Historical precip vs the  non-secondary corrected BCCA projections
precipsummary(obs, "Precip",future,"Precip","Historical Precip vs. Non-secondary corrected BCCA")

#Compare the non-secondary corrected BCCA projections vs the CCFM scaled results
precipsummary(scaledccfm,"scaled",future,"Precip","CCFM vs. Non-secondary corrected BCCA")

#Compare the Historical precip vs the CCFM scaled results
precipsummary(scaledccfm,"scaled",obs,"Precip","CCFM vs. Historical Precip")



```
